### 2023-01-04 RSV_timing_explain/RSV_timing_tutorial.Rmd
add more materials (discussion point with Deus next week?)

---
title: "Estimate RSV onset and peak timing"
subtitle: "Tutorial 2 for transition"
author: 
- Gigi (Zhe Zheng)^[zhe.zheng@yale.edu; zhe.zheng@aya.yale.edu; gigi.zhe.zheng@gmail.com]
- Dan (Daniel Weinberger) 
- Ginny (Virginia Pitzer)
date: "2023-01-04"
fontsize: 10pt
output: 
  beamer_presentation:
  theme: CambridgeUS
  colortheme: "beaver"
  fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

## Background

```{r figure1,echo = F,out.width = "100%", out.extra="page=4"}
knitr::include_graphics(path = "Epidemics_conference.pdf")
```

## Outline

- In section 1, we will first introduce how to find the peak timing of RSV epidemics using harmonic regression (given regular annual/biennial seasonality). We will then talk about identifying the onset of RSV epidemics using second derivative method, regardless of the seasonality of RSV.
\
\
- In section 2, we will learn how to use R to identify peak timing and onset of RSV epidemics.
\
\
- In section 3, we will incorporate the spatial component of RSV epidemics. We will first introduce the concept of spatial autocorrelation and then lean to use R to account for spatial autocorrelation.

## References

Relevant readings:
\
\
- [RSV onset timing at county level](https://doi.org/10.1093/cid/civ331)
\
\
- [RSV peak timing on state level](https://doi.org/10.1371/journal.ppat.1004591)
\
\
- [Comparing RSV onset timing before and during the COVID-19 pandemic](https://doi.org/10.1111/irv.12965)
\
\
- [RSV peak timing at ZIP code level and the drivers of RSV spread](https://doi.org/10.1126/sciadv.abd6421)