### 2023-01-05 RSV_timing_explain/Lecture 11/r examples.Rmd
add Dan and Ginny's class materials

---
title: "Quantitative Methods in Infectious Disease"
subtitle: "Time series examples"
author: "Dan Weinberger"
date: 'November 28, 2018'
runtime: shiny
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(RCurl)
require(shiny)
set.seed(123)
```

## Examples

Here are some worked examples of what we are covering in class

## Generating harmonic curves
-First create an index variable for time (1,2,3...t)
-Then use sine and cosine functions to generate a wave with specified period
-12 month harmonic (sine and cosine waves)
```{r harmonics, echo=TRUE}
t<-1:120
sin12<-sin(2*pi*t/12)
cos12<-cos(2*pi*t/12)
plot(sin12, type='l', col='red')
points(cos12, type='l', lty=2, col='blue')
```

##Vary the period 
```{r period.slider, echo=FALSE}
#library(shiny)
shinyApp(

  ui = fluidPage(
    sliderInput("period", "Period:",
               min=2, max=36, value=12),
    plotOutput("periodPlot")
  ),

  server = function(input, output) {
    output$periodPlot = renderPlot({
      plot(sin12, type='l', col='gray')
      points(sin(2*pi*t/input$period),type='l', col='red', xlab='time' )
      title(paste0('sin(2*pi*t/',input$period,')'))
    },width = "auto", height = "auto")
  }
)
```

#Harmonic Regression

##Generate count data with known periodicity
-120 months total (10 years of data)
-Has both a 10 and 20 month cycle
-Amplitude=1.5
```{r gen_data , echo=TRUE}
n=120
t<-1:n
set.seed(123)
#Generate some  data that has a frequency of 1/10 and 1/20  frequency (period of 10 ,10 )
amp1=0.75
freq1=1/12 # high frequency

amp2=1.5
freq2=1/24 #low frequency

xt1a=amp1*cos(2*3.14159*t*freq1) #component 1
xt1b=amp2*cos(2*3.14159*t*freq2) #component 2
intercept<-0.0
lambda<- exp(intercept+xt1a+xt1b)
xt1=rpois(n,lambda)

```

## View the data we generated
```{r plots1, echo=FALSE}
par(mfrow=c(2,2), mar=c(2,2,1,1))
plot(xt1a, bty='l' , main='Component 1', type='l' , ylim=range(c(xt1a,xt1b)))
plot(xt1b, bty='l' , main='Component 2' , type='l', ylim=range(c(xt1a,xt1b)))
plot(xt1a+xt1b, bty='l' , main='log(Component 1 +Component 2)' , type='l', ylim=range(xt1a+xt1b))
plot(t, xt1,type="l", lwd=3, xaxt="n",bty="l",ylab="Cases of Disease", xlab="Month(n)", main="Cases")
```

##Create the needed harmonic variables with 10, 20, and 40 month periods
```{r harm1, echo=TRUE}
t<-1:120
#Create harmonic variables
sin12=sin(2*pi*t/12)
cos12=cos(2*pi*t/12)
sin24=sin(2*pi*t/24)
cos24=cos(2*pi*t/24)
sin48=sin(2*pi*t/48)
cos48=cos(2*pi*t/48)
```
plot(xt1)

## Fit a simple poisson regression with just 12 month period
```{r harm_reg1}
fit1a <- glm(xt1 ~ sin12+cos12, family='poisson' )
pred1a<- fitted(fit1a)
#summary(fit1a)
par(mfrow=c(1,1))
plot (t, xt1,bty="l", xaxt="n",yaxt="n",xlab="Time",ylab="Cases of Disease", type="p")
points(t,pred1a, type="l",lty=1,col="blue", lwd=3)
```

## Add in 24 month periodicity
```{r harm_reg2}
fit2a <- glm(xt1 ~ sin12+cos12 +sin24 +cos24, family='poisson' )
pred2a<- fitted(fit2a)
#summary(fit2a)
par(mfrow=c(1,1))
plot (t, xt1,bty="l", xaxt="n",yaxt="n",xlab="Time",ylab="Cases of Disease", type="p")
points(t,pred2a, type="l",lty=1,col="blue", lwd=3)
```

## Add in 48 month period 
```{r harm_reg3}
fit3a <- glm(xt1 ~ sin12+cos12 +sin24 +cos24 +sin48 +cos48, family='poisson' )
pred3a<- fitted(fit3a)
#summary(fit3a)
par(mfrow=c(1,1))
plot (t, xt1,bty="l", xaxt="n",yaxt="n",xlab="Time",ylab="Cases of Disease", type="p")
points(t,pred3a, type="l",lty=1,col="blue", lwd=3)
```

## Determine best model with AIC (smaller=better)
Winner is Model 2 (12 and 24 month period). Models 2 and 3 have similar AIC score (within 2 points), so we prefer the simpler model
```{r aic}
 AIC(fit1a)
 AIC(fit2a)
 AIC(fit3a)
```

#Calculate amplitude of the 10 and 20 month periods and their ratio
```{r amp}
beta_sin12<-coef(fit2a)['sin12']
beta_cos12<-coef(fit2a)['cos12']
beta_sin24<-coef(fit2a)['sin24']
beta_cos24<-coef(fit2a)['cos24']

amp12<-sqrt(beta_sin12^2 +beta_cos12^2)
amp24<-sqrt(beta_sin24^2 +beta_cos24^2)
ratio12_24<-amp24/amp12
ratio12_24
amp12 #True value=0.75
amp24 #True value=1.5
```
## log(x+1)=bo+b1sin12+b2cos12+b3sin24+b4cos24

##Fast Fourier Transform
If we don't already know what periods should be present, can we use first use FFT to detect the correct period?
Can sqrt or log stranform data prior to FFT
```{r fft1, echo=TRUE}
# Do a FFT
fft1=fft(sqrt(xt1))
fft1[1:10] 
```
Generates a bunch of imaginary numbers that are hard to interpret

##Function to convert raw power spectrum into something interpretable
```{r fft_func, echo=FALSE}
plot.frequency.spectrum <- function(X.k, xlimits=c(5,n/2)) {
  plot.data  <- cbind(0:(length(X.k)-1), Mod(X.k))

  plot.data[2:length(X.k),2] <- 2*plot.data[2:length(X.k),2] 
  plot.data[2:length(X.k),1] <- n/plot.data[2:length(X.k),1]

  plot(plot.data, t="h", lwd=2,  
       xlab="Period", ylab="Strength", 
       xlim=xlimits, ylim=c(0,max(Mod(plot.data[,2]))), bty='l', main="FFT power spectrum")
}
```

#FFT power spectrum
```{r plot_fft}
par(mfrow=c(1,1))
  plot.frequency.spectrum(fft1, xlimits=c(2,30))

```

##FFT slider
```{r fft.slider, echo=FALSE}
#library(shiny)
shinyApp(

  ui = fluidPage(
    sidebarLayout(position = "left",
    sidebarPanel(
    sliderInput("period1", "Period1:",
               min=2, max=48, value=12),
     sliderInput("amplitude1", "Amplitude1:",
               min=0.2, max=2.5, value=0.75),
    sliderInput("period2", "Period2:",
               min=2, max=48, value=24),
      sliderInput("amplitude2", "Amplitude2:",
               min=0.2, max=2.5, value=1.5),
      sliderInput("int1", "Intercept:",
               min=0, max=5, value=0)
      ),
    mainPanel(
       plotOutput("periodPlot")
    )
  )
  ),

  server = function(input, output) {
    t=1:120
    xt1a=reactive({input$amplitude1*cos(2*3.14159*t/input$period1)}) #component 1
    xt1b=reactive ({input$amplitude2*cos(2*3.14159*t/input$period2)}) #component 2
    intercept<-reactive({input$int1})
    lambda<- reactive ({exp(intercept()+xt1a()+xt1b())})
    xt.slide=reactive ({rpois(n,lambda())})
    fft.s=reactive ({fft(sqrt(xt.slide()))})

    output$periodPlot = renderPlot({
      par(mfrow=c(2,1), mar=c(3,2,1,1))
   plot(t, xt.slide(),type="l", lwd=3, bty="l",ylab="Cases of Disease", xlab="Month(n)", main="Cases")
      plot.frequency.spectrum(fft.s(), xlimits=c(2,max(input$period1, input$period2)))

    },width = "auto", height = "auto")
  })
```

#Wavelets

##Pseudo-RSV data: Simulate time series with 12 and 24 month periods
Imaine this is RSV case data from 2 states, and we want to investigate the
epidemic characteristics in these states and the lag between states
```{r waveset, echo=FALSE}
set.seed(123)
  n=120 # 10 years
  t <- seq(1,n)
 amp1=2.5
 freq1=1/12 # high frequency
 amp2=0.5
 freq2=1/24 #low frequency

 xt1a=amp1*cos(2*3.14159*t*freq1)
 xt1b=amp2*cos(2*3.14159*t*freq2)

#other series shifted by 3 months and with smaller biennial component
 xt2a=amp1*cos(2*3.14159*t*freq1+1)
 xt2b=0.5*amp2*cos(2*3.14159*t*freq2+1)

 #Simulate some poisson count data
 xt1=rpois(n,exp((1+xt1a + xt1b)))
 xt2= rpois(n,exp(2+xt2a+xt2b))
 plot(t, xt1,type="l", lwd=3, xaxt="n",bty="l",ylab="Cases of RSV", xlab="Month(n)", ylim=range(c(xt1,xt2)))
 points(t,xt2, type="l",col="red")
```

```{r wave1, echo=FALSE}
##Source the Wavelet functions from github. Originally from Johannson PLOS Medicine https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1000168
urlX<-getURL("https://raw.githubusercontent.com/weinbergerlab/quant_meth_id/master/WaveletPkg.R")
eval(parse(text = urlX))
#source('WaveletPkg.R')
```

##Run wavelets on the sqrt(series) we generated
```{r wave2, echo=TRUE}
	wave1<- as.data.frame(cbind("x"=(sqrt(xt1)), "t"=t))
	cwt1 <- complete.cwt(wave1, dj=1/10)
	plot.cwt(cwt1)
		wave2 <-as.data.frame(cbind("x"=(sqrt(xt2)), "t"=t)) 
	cwt2 <- complete.cwt(wave2, dj=1/10)
		plot.cwt(cwt2)

```

##Compare phase angle of the 2 series we generated
```{r wave3, echo=FALSE}

	phase1<-phase(cwt1, s=c(10, 14))
	phase2<-phase(cwt2, s=c(10, 14))
	plot(t, phase1,type='l', col="blue", lty=1)
	points(t, phase2,type='l', col="red", lty=1)
```

##Calculate and view the phase difference
```{r wave4, echo=TRUE}
#Take modulo (%%) of phase difference to account for edge effects as described by Grenfell. This corrects for problems at start and begining of cycles
	phase.diff2.1= phase2-phase1
	phasediff2.1_correct <- ((phase.diff2.1 + 3*pi) %% (2*pi)) - (pi)
	#shows corrected phase diff in modulo for the 12 month phase
	plot(t,phasediff2.1_correct,type='l',col="blue", lty=1,ylim=c(-pi,pi))

```

##Convert phase difference to monthhs
```{r wave5, echo=FALSE}
	phase.diff.month <- phasediff2.1_correct * 12/(2*pi)
	plot(t,phase.diff.month,type='l',col="blue", lty=1,ylim=c(-12,12))
```

##Reconstruct series, filtering out noise
```{r wave.recon, echo=FALSE}
 recon1 <- reconstruct.series(cwt1)  #square to get back to original scale
 recon1a <- reconstruct.series(cwt1, sp1=10, sp2=27)^2 #keep periods of 10-27
 recon1b <- reconstruct.series(cwt1, sp1=10, sp2=14)^2 # only keep periods of 10-14
 plot(t,recon1,type='l',col="blue", lty=1)
 points(t,recon1a,type="l",col="red",lty=1)
 points(t,recon1b,type="l",col="green",lty=1)
```

#Compare biennial strength in the 2 series
#Confirms that biennial pattern is stronger in series 1 than series 2
```{r annual.biennial}

biennial.func<-function(ds, period.range.short=c(10,14),period.range.long=c(20,30)){
  annual.indices<-which(ds$periods>=period.range.short[1] & ds$periods<=period.range.short[2])
  biennial.indices<-which(ds$periods>=period.range.long[1] & ds$periods<=period.range.long[2])
  annual.power<-mean(ds$global[annual.indices])
  biennial.power<-mean(ds$global[biennial.indices])
  biennial.annual.ratio<-biennial.power/annual.power
}

series1.biennial.ratio<-biennial.func(ds=cwt1)
series2.biennial.ratio<-biennial.func(ds=cwt2)
series1.biennial.ratio
series2.biennial.ratio
```

##Coherence analysis 
Evaluate coherence between the 2 series. At which periods and which time points are the series in sync?
```{r coherence}
#Coherence analysis
coh1.2 <- complete.coh(wave1,wave2, mc.sim=100, dj=1/10)
plot.coh(coh1.2)
```