### 2023-01-26 5_roadmaps_future_projects/RSV_elderly_vac_cost_eff/Hosp_burden/Hosp_burden_script.Rmd
add materials

---
title: "RSV burden in older adults by age and risk groups"
author: "ZHE ZHENG"
date: "2022-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in time-series data
```{r}
#####################################################################################
# read in dataset and data cleaning
##################################################################################### 
#getwd()
#rm(list = ls())
library(data.table)
library(sas7bdat)

# the time series data will come from SAS output
# we only select the variable that we are interested because the whole data set is huge
# For the cost-effectiveness study, we should add the cost variables 
Wholeresp <- fread("/Users/zhezheng/Box/data/data/recoded_all_states_resp.csv",select = c("agemonth","zip","AGE","HOSPST","resp","resp_prim","flu","flu_prim","rsv","rsv_prim","resp_other","resp_otherprim","bronchio",'bronchio_prim',"pneumo_other","pneumo_otherprim","pneumopneumo","pneumopneumo_prim","pneumosept","pneumosept_prim","asthma","amonthdate","agecat1","riskcat"))

# Depends on the output of the SAS files. fread is faster
# Wholeresp <- read.sas7bdat("/Users/zhezheng/Box/data/data/recoded_all_states_resp.sas7bdat")
```

```{r}
# If interested in analyzing each state, uncomment these codes

# NYresp <- Wholeresp[HOSPST=="NY",]
# AZresp <- Wholeresp[HOSPST=="AZ",]
# FLresp <- Wholeresp[HOSPST=="FL",]
```

```{r}
# include only the complete cases
Wholeresp <- na.omit(Wholeresp)
# recode the date variable
Wholeresp$amonthdate<- as.Date(Wholeresp$amonthdate, "%d%b%Y")

# make sure the age is a numeric variable not character
Wholeresp$AGE <- as.numeric(Wholeresp$AGE)

# Consider using epidemiological calendar
# Wholeresp.1 <- Wholeresp[Wholeresp$amonthdate>"2016-06-01"&Wholeresp$amonthdate<"2021-07-01",]

# sum over individual records to create monthly time-series of RSV cases, flu cases, respiratory hospitalizations by age and risk category 
resp <- Wholeresp[,list(resp = sum(resp),rsv=sum(rsv),flu=sum(flu)), by = 'amonthdate,AGE,riskcat'] # sum over
```

```{r}
# recode age categories to reflect what we interests: "<1","1-<2","2-<5","5-9","10-19","20-59","60-64","65-69","70-74","75-79","80+"

resp$agecat_new[resp$AGE==0]  <- "<1"
resp$agecat_new[resp$AGE==1]  <- "1-<2"
resp$agecat_new[1<resp$AGE&resp$AGE<1]  <- "2-<5"
resp$agecat_new[5<=resp$AGE&resp$AGE<10]  <- "5-<10"
resp$agecat_new[5<=resp$AGE&resp$AGE<10]  <- "5-<10"
resp$agecat_new[10<=resp$AGE&resp$AGE<20]  <- "10-<20"
resp$agecat_new[20<=resp$AGE&resp$AGE<60]  <- "20-<60"
resp$agecat_new[60<=resp$AGE&resp$AGE<65]  <- "60-<65"
resp$agecat_new[65<=resp$AGE&resp$AGE<70]  <- "65-<70"
resp$agecat_new[70<=resp$AGE&resp$AGE<75]  <- "70-<75"
resp$agecat_new[75<=resp$AGE&resp$AGE<80]  <- "75-<80"
resp$agecat_new[80<=resp$AGE]  <- "80+"
```

```{r}
# reshape to create a respiratory hospitalizations array that has a column for time (48 months), a column for age (11 groups) and a column for risk groups (2 groups, may increase. For hierarchical structure, we need at least 3-4 groups to make)
# adjust the first value to reflect the length of time

# test
df2 <- data.frame(x = rep(1:48,22),y = rep(1:2,528), z=rep(1:11,96))

Array1 <- array(data = unlist(df2),
                  dim = c(48, 11, 2),
                dimnames = list(paste0("month",1:48),                              c("<1","1-<2","2-<5","5-9","10-19","20-59","60-64","65-69","70-74","75-79","80+"),
                                c("high risk","low risk")))
```

```{r}
# create a reported RSV array that has a column for time (48 months), a column for age (11 groups) and a column for risk groups (2 groups)

# adjust the first value (48) to reflect the length of time
y <- array(data=NA,dim = c(48,11,2),
           dimnames = list(paste0("month",1:48),                              c("<1","1-<2","2-<5","5-9","10-19","20-59","60-64","65-69","70-74","75-79","80+"),
                                c("high risk","low risk")))

```