### 2023-01-26 5_roadmaps_future_projects/RSV_elderly_vac_cost_eff/RSV_costeffect_elderly.Rmd

---
title: "Project outline: RSV vaccination in older adults"
author: "ZHE ZHENG"
date: "2022-11-16"
output:
  html_document: default
  pdf_document: default
---

### Step 1. Estimate the true burden of hospitalizations across risk and age groups in older adults

##### Background

Risk groups will be defined by ACIP.  
references: [ACIP recommendations](https://www.cdc.gov/vaccines/vpd/pneumo/hcp/who-when-to-vaccinate.html#adults-19-64)

* Low risk population
* High risk population: For adults with any of the conditions or risk factors listed below: Adults who have chronic pulmonary (including asthma), cardiovascular (excluding isolated hypertension), renal, hepatic, neurologic, hematologic, or metabolic disorders (including diabetes mellitus).
Persons who are immunocompromised due to any cause (including but not limited to immunosuppression caused by medications or HIV infection).

##### Estimate the "true" burden of RSV hospitalizations and ED visits
Model structure:\
$Y_{ijk}$ ~ Negative Binomial ($p_{ijk}$,$r$) \
$p_{ijk}=r/(r+λ_{ijk})$ \
where $Y_{ijk}$ denotes the number of all-cause respiratory hospitalizations at time (month) $i$, in age group $j$, and risk group $k$. 
$i = 1,2,...,48$ \
We will analyze hospitalizations in 11 age categories (<1, 1, 2-4, 5-9, 10-19, 20-59, 60-64, 65-69, 70-74, 75-79, 80+ years).
$j = 1,2,...,12$ 
\
\
We define the expected value as a function of covariates and random effects such that:

$$\lambda_{ijk}=\beta_{0jk}+\alpha_{1g(i)}+\alpha_{2m(i)}+\beta_{1jk}RSV_{i}+\beta_{2g(i)jk}Flu_{ik}$$
<span style="color:red">  $RSV_{i}$ </span> (means every risk group has the same RSV timing)
\
\
The number of RSV hospitalizations/ED visits in each age group and risk will be <span style="color:blue">$\beta_{1jk}RSV_{i}$</span>

##### Estimate death rate in each risk group

(1) Include DIED element in SAS code
(2) Calculate the death rate in each risk and age group and assume that the death rate is the same between reported cases and the cases that did not reported

\
\

### Step 2.  Fitting transmission model to time-series output

Parameters to estimate (in total 6?):

(1) Transmission parameter (related to transmission probability)
(2) Amplitude of seasonality
(3) Timing of seasonality
(4) Duration of maternal immunity
(5) Proportion of third LRIs that are hospitalized in
older adults (if we assume their contact patterns are similar, we only need to estimate two parameters)
* 60-64 (low $H_{60l}$ and high risk $\frac{\beta_{60h}}{\beta_{60l}} H_{60l}$)
* 65-69 (low $\frac{\beta_{65l}}{\beta_{60l}}H_{60l}$ and high risk $\frac{\beta_{65h}}{\beta_{60l}}H_{60l}$)
* 70-74 (low $\frac{\beta_{70l}}{\beta_{60l}}H_{60l}$ and high risk $\frac{\beta_{70h}}{\beta_{60l}}H_{60l}$)
* 75-79 (low $\frac{\beta_{75l}}{\beta_{60l}}H_{60l}$ and high risk $\frac{\beta_{75h}}{\beta_{60l}}H_{60l}$)
* 80+ years (low $H_{80l}$ and high risk $\frac{\beta_{80h}}{\beta_{80l}}H_{80l}$)
\
\

### Step 3.  Vaccine efficacy

References:

(1) [Efficacy on infectiousness,preF vaccine](https://www.nejm.org/doi/full/10.1056/NEJMoa2116154)
(2) [Efficacy on infectiousness, vector vaccine](https://academic.oup.com/jid/article/226/3/396/6064820)
(3) [Efficacy on symptomatic infection, Pfizer](https://www.pfizer.com/news/press-release/press-release-detail/pfizer-announces-positive-top-line-data-phase-3-trial-older)
(4) [Efficacy on symptomatic infection, GSK](https://www.gsk.com/en-gb/media/press-releases/gsk-s-older-adult-respiratory-syncytial-virus-rsv-vaccine-candidate/)
(5) [Efficacy on symptomatic infection, Janssen](https://www.jnj.com/janssen-announces-phase-2b-data-demonstrating-its-investigational-rsv-adult-vaccine-provided-80-protection-against-lower-respiratory-infections-in-older-adults)

vaccine effects:

(1) Reduce the risk of infection? $(1-VE_{s})\sigma_3\lambda$

\ where $VE_{s}$ is the vaccine efficacy against infection, $\sigma_3$ is the relative susceptibility during subsequent infections, and $\lambda$ is the force of transmission.

(2) Reduce the infectiousness of vaccinees?
$(1-VE_{i})I_v$

\ where $VE_{i}$ is the vaccine efficacy to reduce infectiousness and $I_v$ is the number of vaccinated individuals who are infected by wild-type RSV

(3) Duration of protection? Assume 1 year as a start and sensitivity analysis of 2 years and 5 years

(4) In later stage: we can add age related immunogenicity?
\
\

### Step 4.  Economic Model

###### Health benefits and related costs
(1) Treatment outcomes and costs
(2) QALY estimation
(3) Vaccination costs

* Sample from the uncertainty distributions of the model parameters while running the model 1,000 times to generate 95% prediction interval for the number of infections, GP, ED, hospitalizations, and deaths averted

* Calculate the average net monetary benefit of each vaccination strategy (and no vaccination) for WTP values varying from $10,000 to $50,000 per QALY averted (in increments of $100).

* Plot the Cost-Effectiveness Acceptability Frontier