### 2023-01-26 4_RSV_disease_burden/4_RSV_disease_burden.Rmd
UPDATE

---
title: "Bayesian model in JAGS"
subtitle: "Tutorial 3 for transition"
author: "Zhe Zheng (Gigi)^[zhe.zheng@yale.edu]"
date: "2023-01-16"
fontsize: 10pt
output: 
  beamer_presentation:
  theme: CambridgeUS
  colortheme: "beaver"
  fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline

- Section 1: Background knowledge of estimating disease burden (hospitalizations/mortality etc) with statistical models.
\
\
- Section 2: Background knowledge of Hierarchical Bayesian regression.
\
\
- Section 3: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations.
\
\
- Section 4: R code for running the Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups.

## Section 1: Background knowledge of estimating disease burden with statistical models.

There are 5 main statistical methods to estimate the hospitalization/mortality attributable to respiratory virus infection. In Dan's class, Public Health Surveillance, he teaches all four methods in details. Here, we will only give a brief introduction to each of the method and provide the link to the initial publications.

1. **Serfling regression**

2. **Periseason differences**

3. **Poisson regression with log link**

4. **Negative binomial regression with identity link**

5. **Box-Jenkins transfer function**

## Initial publications of the method

1. *Serfling regression*
  - [Methods for Current Statistical Analysis of Excess Pneumonia-influenza Deaths](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1915276/pdf/pubhealthreporig00078-0040.pdf)
  - [The impact of influenza epidemics on mortality: introducing a severity index](https://ajph.aphapublications.org/doi/epdf/10.2105/AJPH.87.12.1944)
  - [Impact of Influenza Vaccination on Seasonal Mortality in the US Elderly Population](https://jamanetwork.com/journals/jamainternalmedicine/article-abstract/486407)

2. *Periseason differences*
  - [The effect of influenza on hospitalizations, outpatient visits, and courses of antibiotics in children](https://www.nejm.org/doi/full/10.1056/nejm200001273420401)
  - [Respiratory illness associated with influenza and respiratory syncytial virus infection](https://adc.bmj.com/content/archdischild/90/7/741.full.pdf)
  - [Influenza and the Rates of Hospitalization for Respiratory Disease among Infants and Young Children](https://www.nejm.org/doi/full/10.1056/nejm200001273420402)

## Initial publications of the method (continued)

3. *Poisson regression with log link*
  - ()

4. *Negative binomial regression with identity link*
  - ()

5. *Box-Jenkins transfer function*
  - ()

## Other useful references

- []

- [The Need for Validation of Statistical Methods for Estimating Respiratory Virus–Attributable Hospitalization](https://academic.oup.com/aje/article/170/7/925/91526)